### 2015-12-08 paper/paper.tex
Completed moving paper.txt to paper.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[toc,page]{appendix}

%opening
\title{Tools and Ideas for the Proceedural Generation of Shadow Sculptures.}
\author{James Hoctor, Chris Chappell}

\begin{document}

\maketitle

\begin{abstract}
This paper discusses the creation of three dimensional objects,
given a shadow, a light source, and a wall, such that the object
creates the shadow on the wall when placed under the light source.
This problem was inspired by the artwork of Larry Kagan, a retired
professor at RPI. The use of classical and real valued genetic
algorithms to create these objects is demonstrated. A visualization
tool is described that shows how the object would look as a finished
product. We describe our respresentation of an object which is based
on the techniques Kagan uses in his art. An efficient way to
calculate the shadow cast by an object with this representation is
required in order to evaluate fitness in the genetic algorithm. We
describe our approach to shadow rendering. One component of fitness
in the genetic algorithm is the fidelity of the cast shadow to the
desired result. We use an industry standard algorithm which is often
applied in image quality analysis to judge the quality of the
shadows. Having described the means to evaluate the fitness of a
candidate object, we proceed to explain the recombination techniques
we used to create each generation of objects from the parent generation.
\end{abstract}

\section*{Introduction}
This paper addresses the proceedural generation of artwork, if such
a thing is even possible. We wish to create art in the style of
Larry Kagan, whose medium is metal and the shadows it casts. In a
typical Kagan piece there is a metal object made from steel bars
that have been bent to shape and welded together to form a whole.
This assembly is mounted on a wall (painted white) and a spot light
shines on it to cast a shadow on the wall that forms an image. In a
naive sense, the problem  of creating something like Kagan's work is
a very simple one: create a metal object in the shape of the desired
shadow but scaled down by half, then mount this half way between the
light source and the wall. The mounting can be done by means of
straight bars that travel in the shadow from the rest of the object
to the wall. But this is not the kind of result we desire. A major
challenge of this project is determining the other factors that make
a Kagan piece interesting beyond the mechanical fact that an image
is present in the form of a shadow. In particular, the metal portion
of a genuine Kagan is often visually chaotic and does not contain
recognizable forms. (Exceptions to this are discussed in the section
\ref{kagan}.) Our approach to generating artwork in this style
is therefore to use a genetic algorithm to search for a configuration
of metal that meets a variety of conditions among which is the
obvious requirement that the shadow appear as desired.

Other components of fitness of the object in the genetic algorithm
are its structural stability, its visual chaoticness, and its ability
to be fabricated using techniques similar to what Kagan uses. We did
not fully address structural stability in our research. Visual
chaoticness is achieved by the random nature of the selection process.
Fabricability is included because it adds another contraint in a
problem that has few. It is was key consideration when we decided
how to represent these sculptures.

\section{Larry Kagan's Work} \label{kagan}
In researching this problem we made a brief study of Kagan's work for
several reasons. Firstly, we wanted to identify a subset of his work
that might be most easily imitated. Secondly, our analysis shows how
necessary this restriction is. Kagan's work is too varied for the
whole body of work to be imitated. There is also the question of how
Kagan selects the image he wishes to create from shadow. In all of our
work we leave this crucial piece of creative heavy lifting to the
experts: humans.

Each of Larry Kagan's sculptures consists of a metal structure
composed of bent rods of metal mounted to a wall and illuminated by
a bright light from a carefully chosen location. The eye is drawn to
the shadow produced this way, which forms a familiar image, while the
rods themselves often have no obvious pattern. In the simplest case,
the shadow image is a line drawing and the rods are attached to the
wall at several points. For example, the piece "Frankfurt Chair" is
like this. Kagan has created many variations on this idea including
incorporating some of the bent rods in the recognizable image while
most of the image is formed by the shadow, as seen in "Beach Chair".
"Intersection of Two Circles" shows that a sculpture composed of rods
can have a solid shadow with a geometric shape. The light source can
be on the ceiling cented over the piece or be off to one side as it
is in "At". With the light source on one side the shadow falls
predominantly on the side of the sculpture. Despite the fact that a
single small bright light source is installed to cast the shadow for
each sculpture, soft egdes of the shadow are noticeable in many of
these works. In "Mosquito I", Kagan uses this effect in a dramatic
way; the image present in the shadow of this sculpture is a mosquito
with its own shadow cast by an imaginary light source. In order to
make the distinction obvious, Kagan allowed soft shadow edges to show
in the portion of the image corresponding to the imaginary shadow of
the mosquito, but created the image of the mosquito with dark
hard-edged shadows.

\section{How We Represent Sculptures and Why}
There are two different representations of sculptures used in this project. Both were chosen to make it likely that a sculpure that can be represented in our programs will be possible to make using techniques similar to Kagan's.

The first representation is used when saving a sculpure as a digital file. These files can be given as input to the visualization system that is described later in this paper. The informal specification of this format can be found in the file 'ad file spec.txt' in the git repository submitted with this paper. An example file in this format can be generated by the script make-art-description.py, which is also in the git repository.

The contents of such a file include three things that represent the sculpture: the position of the light source (actually a list of point sources, so we can simulate a diffuse source), the radius of the bar stock to use in constructing the sculpture and a list of 'strands'. Each strand is a list of points in 3D space, and it represents a space-filling shape composed of the union of spheres centered at each point and of cylinders with the centers of their end faces at consecutive points in the strand. The idea is that a strand approximates a smoothly bending piece of steel bar. One could construct the sculpture by making each strand from bar stock, then holding each strand in place and welding them together wherever they make contact.

This representation lends itself to checking whether the sculpture can be feasibly made or not. If any single strand cannot be made then the whole cannot. A strand might be impossible to make because it bends too sharply, or self intersects. The whole assembly might also fail to be fabricable if two strands intersect more than the small amount of intersection that would result from a weld. For comparison, one example of a representation that cannot be so easily checked for fabricability is a mesh of points describing the surface of the sculpture. This representation does lend itself to calculating shadows however.

The second representation that we use also makes use of the idea of spheres connected by cylinders. In fact, the representation is a simple graph with nodes representing the centers of spheres and edges representing cylinders connecting the centers of two spheres. This is the representation we propose using for the genetic algorithm.

Since any sculpture having a representation of the first kind has a representation of this second kind, but not vice versa, it is more likely that a sculpture with a representation of this second type will fail to be fabricable. Because of this, we propose that this representation be used with several checks. For example, one might check that no node has an order greater than 4 or 5, and that at least a certain percent of nodes have an order of 2 or fewer.

\section{How We Render Sculptures}
Using the method of spheres and cylinders, objects can be graphically modeled and displayed in 3D rendering systems fairly easily.  One method of doing this that we tested was using OpenGL to build a little engine that could render the object and a small surface such that the object and the shadow could be viewed from multiple angles, as it would look if it were printed.  The shadow could be imported as a texture, using OpenGL's \texttt{GL\_TEXTURE2D} object, and then fitted to the surface that would compose the wall.  This would allow modularity and would separate the methods that composed the creation of the shadow and its rendering.  While this may not be optimal, it is the most easily supported method using OpenGL legacy code.

For a more updated model that could possibly fix both the issue of slow shadow rendering time and separation of the shadow rendering from the object's rendering, one may instead of legacy OpenGL use OpenGL 3.3 or above, which adds the use of fragment shaders, and can render shadows from point light sources fairly easily.  The question that arises in this method, however, is how one would compare two shadows to determine accuracy and fitness of an object.  This would require further research and testing.

There are many video game engines that have builtin methods for shadow generation, with many options and customizable settings for doing so.  This may be the simplest option for rendering objects, and possibly even would also allow for comparison and heuristic generation.  Such engines would include Unity, Unreal Engine, and Source 2, just to name a few.  As our knowledge did not extend to the realm of writing video games, we felt more comfortable working with OpenGL.

\section{How We Render Shadows}
We investigated two CPU based ways to render the shadows of objects with a representation of the first kind (a list of strands). We believe that the increased speed of a GPU-based approach would be beneficial to the implementation of a genetic algorithm.

The first method is partially implemented in the file 'rayTracing.py' on the branch 'master' in the submitted git repository. In this method the cast shadow of each sphere and cylinder is computed individually and merged. Here is some pseudo-code for rendering the shadow of a collection of spheres and cylinders illuminated by a single point light source:

input: l = position of light
	   r = the radius of each sphere and cylinder
	   spheres = a list of positions of centers of spheres
	   cylinders = a list of pairs of positions specifying the end faces of the
					cylinders
output: pixels = array of black and white pixels indicating on which areas of the
					wall light is incident

initialize pixels to be uniformly white (illuminated)

for each sphere in spheres:
	if the sphere is entirely above the light or below the shadow plane:
		it casts no shadow; continue to the next sphere
	if the top of the sphere is beneath the light and the bottom is above the plane:
		The shadow cast on the plane will be elliptical.
		Calculate the equation of the cone formed by all lines passing through the
				light source and tangent to the sphere.
		Find the equation of the ellipse by intersecting this cone with the shadow
				plane.
		The equation has the form $A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0$.
		Rearrange the equation to have the form $G(x)*y^2 + H(x)*y + I(x) = 0$, where
				$G$, $H$, and $I$ are 2nd order or lower polynomials in $x$.
		For each column $c$ in pixels:
			Find the value of $x$ that corresponds to $c$.
			Calculate $G(x)$, $H(x)$, and $I(x)$.
			Use the quadratic formula to find two values of $y$ satisfying the
					equation for the boundary of the ellipse.
			Convert the values of y to row coordinates in the image.
			In column $c$ and all rows $r$ between the two calculated:
				color pixels[r][c] black
	if the sphere intersects the shadow plane:
		Handling this case is at least as complicated as handling a sphere strictly
				above the plane.
		The circular intersection of the circle and the plane is shadowed.
		Any additional shadowing is a subset of the elliptical shadow described in
				the previous case.
		This case is not handled by our code.
	if the sphere intersects the light:
		Color all of pixels black.
		Return from the proceedure early.
		This case is not handled by our code.
	if the top of the sphere is above the light, its bottom below the light, without
			intersecting it:
		The shadow cast is hyperbolic.
		This case is not handled by our code, but it would likely be similar to
				the case of elliptical shadows.

for each cylinder in cylinders:
	//similar to the above but cylinders cast different shaped shadows
	// ...
	This case is not handled by our code.

This proceedure needs to be repeated for each point light source, creating a separate image for each. Then these images need to be averaged, pixel by pixel. When generating a 1024x720 image of the shadow this averaging process took about 1.2 seconds for our python implementation. With a 3000x3000 image it took almost 13 seconds. It should be possible to take advantage of data parallelism here and do the averaging process in OpenGL on a GPU, or perhaps with single instruction multiple data technologies like SSE and AVX.

We did not implement casting the shadows of cylinders by this method, but there is an optimization that deserves mention. It results from the fact that we can assume there is a sphere present at each end of any cylinder. Consequentially, any cylinder that is aligned with one face toward the light source casts a shadow only where a sphere has already cast a shadow. The optimization that arises from this is to ignore cylinders with such an orientation in the loop over all cylinders.

Furthermore, even when the cylinder is not aligned this way we can skip drawing the portion of the shadow that results from the end faces, for the same reason. In this case, the shadow drawn for a cylinder would be a quadrilateral. Imagine two planes, each of which is tangent to the circular wall of the cylinder and passes through the point light source. They are tangent on opposite sides of the cylinder. Each plane has an intersection point with each circular face of the cylinder. The four such intersection points are co-planer, and the quadrilateral they form is entirely inside the cylinder. The shadow of this quadrilateral is thus a part of the shadow of the cylinder. The remaining part of the cylinder's shadow should also be shadowed by the spheres.

Appendix \ref{a} presents a derivation of the equation of the elliptical boundary of the shadow of a sphere in the most typical case.

In an attempt to render shadows more quickly, another technique was attempted. The code for this technique can be found in the newRayTracing branch of the git repository. This second method is a refinement of the following naive method:

Suppose you have a pair of proceedures that are used to efficiently check whether a line segment intersects a sphere or a cylinder. You can iterate over the lights, and over all the pixels in the image, and construct the line segment from the light to the center of the pixel. Then go through all the spheres and cylinders checking for an intersection. At the first intersection, the light intensity at the pixel is decremented and we move on to the next pair of a light and a pixel. The motivation behind this method is that there is no step of merging the shadow of each point light source. [Consider writing out pseudo code for this.]

The innermost loop (over all spheres and cylinders) can be accelerated by use of a space partitioning method.

The refinement that can be added is simplest to understand if we restrict ourselves temporarily to intersections with spheres. For each light source we imagine a pyramid with the region-of-interest of the shadow plane as its base and the light source as its vertex. We divide the pyramid into chunks and to each chunk associate all the spheres with a center near the chunk. When checking if a line segment inersects any sphere, we can ignore any sphere not in a chunk near the line segment. OK, but how do we organize this in practice? In our implementation, the pyramid is first divided into horizontal regions...

[Mention inspiration from nearest neighbor algorithm]

It should also be very fast to render shadows using openGL. Shadowmaps are a standard technique for rendering dynamic shadows. In shadowmapping, a scene is rendered from the perspective of a point light source and the resulting "image" is colored according to how far each part of the image is from the light. Shadowmapping typically goes on to use this information to shade a rendering of the same scene from another perspective, but we only want an image of the shadow on the wall.

\section{Genetic Algorithm Solutions}
We believe approximate solutions to the forward problem may be possible using genetic algorithms for creating "abstract" data that fits a given shadow.  The question then becomes how to represent data for each possible solution to the problem in a genome that allows simple crossovers and mutations.  Some possible representations are as follows:

The simplest method of representing nodes and edges would be to keep the number of both constant, moving the nodes around while retaining the same edge mappings for all genomes.  This way, one could represent each node as an unsigned integer, or an array of three unsigned integers representing the positive incremental distance from some origin point.  Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.  This allows the bounding of the entire system in a cube, which we can use to define our 0 and our \texttt{MAX\_INT} for each coordinate direction.  Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.  This would be the genome for the genetic algorithm.  The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the methods above.  Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.  Since the nature of the problem allows for many local solutions, the ability of such a genetic algorithm to converge may be slowed.

One may increase the variability of the genetic algorithm by introducing edge mapping to the genome.  One may represent an edge between node A and node B as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.  This allows for easy genetic combinations, with the ability to vary the number of edges in the graph.  The issue that this does not take care of is the connectivity of the graph.  One would have to either add a penalty to the fitness function which performed a depth-first search on the nodes and penalized unconnected graphs, or alternatively, one could modify the base constructs of the genetic algorithm, making it continue to perform attempts at combining two genomes until it had children that were all fully connected.  For the first option, the question would become how strongly one would decrease the fitness of an unconnected graph, and whether the number of unconnected pieces should increase that penalty.  The second option would allow us to use the same fitness function, but may add extra levels of complexity to the problem due to the required repeated attemps within single generations.

The last important part of the genetic algorithm is a method of performing mutations on genomic data.  Here are some options in consideration:

By simply mutating the data in the genome, one would cause one or multiple nodes to be moved to a new location, possibly on the same plane, but possibly not.  This would be the simplest method of performing a mutation, but depending on how the nodes are laid out, it may mutate a node outside of the acceptable region given for nodes in the given problem.  Whether this is allowed or not depends on the situational requirements given.

Another mutation would be to swap the position of two nodes.  This would not require any checking to see whether any node is outside the mapped region, but it would not add anything new to the domain of a generation from the previous generation, unless there was little or no variation of edges in the whole generation.

Mutations can also be performed on edge data.  This can simply be done by inverting bits in the adjacency matrix, which would cause edges to disappear or new edges to occur.  This could be used along with the first mutation option to give the complete range of options for convergence, barring adding or removing nodes themselves.  The downside to this form of mutation is that it could make the graph unconnected, which, like doing combinations on edges, would require either some penalty in the fitness function, or a checker to make sure any edge that is removed is not required to keep the graph fully connected.

\section{Future Work}
What is the best way to use point sources to simulate a diffuse light source? Is there a better way than point light sources to simulate a diffuse source?

How can the partitioning of space in the second rendering method be improved to allow faster shadow drawing? More equally sized partitions? Adaptively sized partitions?

The inverse square law is not respected by our shadow drawing algorithms. (Does this Include the OpenGL one?) Can this law be incorporated into our algorithms? Is it even necessary to include this or does the human visual system ignore this variation in brightness? Mention the image processing technique histogram equalization.

The visualization tool lacks the ability to draw shadows cast on the object by itself. These shadows are not an important component of Kagan's art as far as we can tell.

Is it a good idea to combine the two cpu-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?

\begin{appendices}
\chapter{A} \label{a}
In this appendix we will show in detail the calculation of the boundary of a shadow cast by a sphere by a point light source. The shadow is cast on the z=0 plane. On the z-axis the sphere is strictly between the light and the plane. We are given the location of the 
\end{appendices}

\end{document}

### 2015-12-09 rayTracing.py
Prepare rayTracing.py for additional code to render the shadows of cylinders.

@@ -6,14 +6,17 @@ from time import time
 
 # convert a point from floating point coordinates to image coordinates
 def convert_point(pt):
-    x = int( (pt[0] - x0)*imagex/(x1-x0) )
-    y = int( (y1 - pt[1])*imagey/(y1-y0) )
+    x = int(round( (pt[0] - x0)*imagex/(x1-x0) ))
+    y = int(round( (y1 - pt[1])*imagey/(y1-y0) ))
     if 0 <= x < imagex and 0 <= y < imagey:
         return (x,y)
     else:
         return None
 
-def cast_shadow(pix, light, point, radius, region, resolution):
+def cast_shadow_cylinder(pix, light, point1, point2, radius, region, resolution):
+    pass
+
+def cast_shadow_sphere(pix, light, point, radius, region, resolution):
     x0,x1,y0,y1 = region
     imagex,imagey = resolution
     if point[2]+radius >= light[2]: #this is the case where a hyperbolic shadow
@@ -72,7 +75,11 @@ def main(infile, outfile):
         pix = im.load()
         for strand in strands:
             for point in strand:
-                cast_shadow(pix, light, point, radius, region, resolution)
+                cast_shadow_sphere(pix, light, point, radius, region, resolution)
+            for i in xrange(len(strand)-1):
+                point1 = strand[i]
+                point2 = strand[i+1]
+                cast_shadow_cylinder(pix, light, point1, point2, radius, region, resolution)
         ims.append(im)
     
     t2 = time()

### 2015-12-09 paper/paper.tex
Changes to Appendix in the LaTeX document.

@@
 \chapter{A} \label{a}
+
 In this appendix we will show in detail the calculation of the boundary of a shadow cast by a sphere by a point light source.
@@
 On the z-axis the sphere is strictly between the light and the plane.
-We are given the location of the
+We are given $\vec{l}$, the location of the point light source, $\vec{c}$, the center of the sphere, and $r$, its radius.
+
+The first step is to derive the equation of the shadow cone.
+A generic point on the cone will be denoted by $\vec{x}$.
+The axis of the cone passes through $\vec{l}$ and through $\vec{c}$.
+The defining property of $\vec{x}$ is
+\begin{equation}\label{xproperty}
+\|orth_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| \cdot k
+\end{equation}
+where $k$ is a constant characteristic of the aperture of the cone.
+Equation $(\ref{xproperty})$ says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
+The value of $k$ can be calculated from any point on the cone.
+We will calculate $k$ from the projection and orthogonal projection of a point of tangency, $\vec{p}$, of the sphere and cone.
+
+The vector $\vec{p} - \vec{l}$ is along a generatrix of the cone, so it is perpendicular to $\vec{p} - \vec{c}$, which points from the center of the sphere to the point of tangency.
+Therefore, a right triangle is formed by $\vec{p}$, $\vec{c}$, and $\vec{l}$, with $\vec{c}$ and $\vec{l}$ at the ends of the hypotenuse.
+The length $L=\|\vec{c}-\vec{l}\|$ can be calculated from the givens and the distance $r=\|\vec{p}-\vec{c}\|$ is a given.
+By the pythagorean theorem the length of the other leg is $L'=\|\vec{p}-\vec{l}\|=\sqrt{L^2-r^2}$.
 \end{appendices}

### 2015-12-09 paper/paper.tex
Maybe halfway done with this appendix.

@@
 \usepackage[toc,page]{appendix}
+\usepackage{amsmath, amsthm, amssymb}
 
@@
 In this appendix we will show in detail the calculation of the boundary of a shadow cast by a sphere by a point light source.
-The shadow is cast on the z=0 plane.
+The shadow is cast on the $z=0$ plane.
 On the z-axis the sphere is strictly between the light and the plane.
@@
 where $k$ is a constant characteristic of the aperture of the cone.
-Equation $(\ref{xproperty})$ says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
+Equation (\ref{xproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
 The value of $k$ can be calculated from any point on the cone.
@@
 
-The vector $\vec{p} - \vec{l}$ is along a generatrix of the cone, so it is perpendicular to $\vec{p} - \vec{c}$, which points from the center of the sphere to the point of tangency.
+The vector $\vec{p} - \vec{l}$ is along a generatrix of the cone, so it is perpendicular to $\vec{p} - \vec{c}$, which points from the center of the sphere to the point of tangency (figure [reference to "sphere shadow figure 1.png"]).
 Therefore, a right triangle is formed by $\vec{p}$, $\vec{c}$, and $\vec{l}$, with $\vec{c}$ and $\vec{l}$ at the ends of the hypotenuse.
@@
 By the pythagorean theorem the length of the other leg is $L'=\|\vec{p}-\vec{l}\|=\sqrt{L^2-r^2}$.
+In figure [reference to "sphere shadow figure 2.png"] the triangle formed by the projection and orthogonal projection of $\vec{p}-\vec{l}$ onto $\vec{c}-\vec{l}$ is shown with the triangle formed by $\vec{p}$, $\vec{c}$, and $\vec{l}$.
+The triangles are similar because they share two congruent angles.
+Thus their sides have the same proportions.
+In particular:
+\[
+k = {\|orth_{\vec{c}-\vec{l}} (\vec{p} - \vec{l})\|\over \|proj_{\vec{c}-\vec{l}} (\vec{p} - \vec{l})\|} = {r'\over L''} = {r\over L'} = {r\over\sqrt{L^2-r^2}}
+\]
+Now we use the defintions of $orth_{\vec{u}} \vec{v}$ and $proj_{\vec{u}} \vec{v}$ to simplify equation (\ref{xproperty}).
+\begin{align*}
+proj_{\vec{u}} \vec{v} = \vec{u} {\vec{u}\cdot\vec{v}\over \vec{u}\cdot\vec{u}}\\
+orth_{\vec{u}} \vec{v} = \vec{v} - proj_{\vec{u}} \vec{v}
+\end{align*}
+Start by squaring both sides of \ref{xproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $y = \vec{x}-\vec{l}$.
+\begin{align*}
+\|\vec{y} - \vec{d} {\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}}\|^2 = \|\vec{d} {\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
+\vec{y}\cdot\vec{y}-2\vec{y}\cdot\vec{d} {\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}({\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}})^2 = (\vec{d}\cdot\vec{d}) ({\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}})^2 \cdot k^2
+\end{align*}
+Multiply each side by $\vec{d}\cdot\vec{d}$:
+\begin{align*}
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{y})^2+(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2\\
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2
+\end{align*}
+Now we find the intersection of this cone with the $z=0$ plane.
+
 \end{appendices}

### 2015-12-14 paper/paper.tex
something something did something

--- 2015-12-08 paper/paper.tex
@@
-\documentclass[a4paper,10pt]{article}
+\documentclass[letter,10pt]{article}
 \usepackage[utf8]{inputenc}
@@
 the shadow image is a line drawing and the rods are attached to the
-wall at several points. For example, the piece "Frankfurt Chair" is
+wall at several points. For example, the piece ``Frankfurt Chair'' is
 like this. Kagan has created many variations on this idea including
 incorporating some of the bent rods in the recognizable image while
-most of the image is formed by the shadow, as seen in "Beach Chair".
-"Intersection of Two Circles" shows that a sculpture composed of rods
+most of the image is formed by the shadow, as seen in ``Beach Chair''.
+``Intersection of Two Circles'' shows that a sculpture composed of rods
 can have a solid shadow with a geometric shape. The light source can
 be on the ceiling cented over the piece or be off to one side as it
-is in "At". With the light source on one side the shadow falls
+is in ``At''. With the light source on one side the shadow falls
 predominantly on the side of the sculpture. Despite the fact that a
@@
 each sculpture, soft egdes of the shadow are noticeable in many of
-these works. In "Mosquito I", Kagan uses this effect in a dramatic
+these works. In ``Mosquito I'', Kagan uses this effect in a dramatic
 way; the image present in the shadow of this sculpture is a mosquito
@@
 These files can be given as input to the visualization system that is described later in this paper.
-The informal specification of this format can be found in the file 'ad file spec.txt' in the git repository submitted with this paper.
+The informal specification of this format can be found in the file `ad file spec.txt' in the git repository submitted with this paper.
 An example file in this format can be generated by the script make-art-description.py, which is also in the git repository.
 
-The contents of such a file include three things that represent the sculpture: the position of the light source (actually a list of point sources, so we can simulate a diffuse source), the radius of the bar stock to use in constructing the sculpture and a list of 'strands'.
+The contents of such a file include three things that represent the sculpture: the position of the light source (actually a list of point sources, so we can simulate a diffuse source), the radius of the bar stock to use in constructing the sculpture and a list of `strands'.
 Each strand is a list of points in 3D space, and it represents a space-filling shape composed of the union of spheres centered at each point and of cylinders with the centers of their end faces at consecutive points in the strand.
@@
 
-The first method is partially implemented in the file 'rayTracing.py' on the branch 'master' in the submitted git repository.
+The first method is partially implemented in the file `rayTracing.py' on the branch `master' in the submitted git repository.
 In this method the cast shadow of each sphere and cylinder is computed individually and merged.
@@
 Shadowmaps are a standard technique for rendering dynamic shadows.
-In shadowmapping, a scene is rendered from the perspective of a point light source and the resulting "image" is colored according to how far each part of the image is from the light.
+In shadowmapping, a scene is rendered from the perspective of a point light source and the resulting ``image'' is colored according to how far each part of the image is from the light.
 Shadowmapping typically goes on to use this information to shade a rendering of the same scene from another perspective, but we only want an image of the shadow on the wall.
@@
 \section{Genetic Algorithm Solutions}
-We believe approximate solutions to the forward problem may be possible using genetic algorithms for creating "abstract" data that fits a given shadow.
+We believe approximate solutions to the forward problem may be possible using genetic algorithms for creating ``abstract'' data that fits a given shadow.
 The question then becomes how to represent data for each possible solution to the problem in a genome that allows simple crossovers and mutations.
@@
 \begin{appendices}
-\chapter{A} \label{a}
+\chapter{Appendix A} \label{a}
+
 In this appendix we will show in detail the calculation of the boundary of a shadow cast by a sphere by a point light source.
@@
 On the z-axis the sphere is strictly between the light and the plane.
-We are given the location of the
+We are given $\vec{l}$, the location of the point light source, $\vec{c}$, the center of the sphere, and $r$, its radius.
+
+The first step is to derive the equation of the shadow cone.
+A generic point on the cone will be denoted by $\vec{x}$.
+The axis of the cone passes through $\vec{l}$ and through $\vec{c}$.
+The defining property of a point $\vec{x}$ is
+
+$\|proj_{\vec{p}-\vec{l}} \vec{x}\| = \|orth_{\vec{p}-\vec{l}} \vec{x}\|c$
+
+where c is a constant characteristic of the aperture of the cone.
+
 \end{appendices}

### 2015-12-14 paper/paper.tex
Added figures to the appendix.
Still need to put them in line with the text instead of on the next page.

@@
 \usepackage{amsmath, amsthm, amssymb}
+\usepackage{graphicx}
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
 
@@
 
-The vector $\vec{p} - \vec{l}$ is along a generatrix of the cone, so it is perpendicular to $\vec{p} - \vec{c}$, which points from the center of the sphere to the point of tangency (figure [reference to "sphere shadow figure 1.png"]).
+The vector $\vec{p} - \vec{l}$ is along a generatrix of the cone, so it is perpendicular to $\vec{p} - \vec{c}$, which points from the center of the sphere to the point of tangency (see figure \ref{fig:shadow_1}).
 Therefore, a right triangle is formed by $\vec{p}$, $\vec{c}$, and $\vec{l}$, with $\vec{c}$ and $\vec{l}$ at the ends of the hypotenuse.
@@
 By the pythagorean theorem the length of the other leg is $L'=\|\vec{p}-\vec{l}\|=\sqrt{L^2-r^2}$.
-In figure [reference to "sphere shadow figure 2.png"] the triangle formed by the projection and orthogonal projection of $\vec{p}-\vec{l}$ onto $\vec{c}-\vec{l}$ is shown with the triangle formed by $\vec{p}$, $\vec{c}$, and $\vec{l}$.
+In figure \ref{fig:shadow_2} the triangle formed by the projection and orthogonal projection of $\vec{p}-\vec{l}$ onto $\vec{c}-\vec{l}$ is shown with the triangle formed by $\vec{p}$, $\vec{c}$, and $\vec{l}$.
 The triangles are similar because they share two congruent angles.
@@
 \]
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_1.png}
+    \caption{Awesome Image}
+    \label{fig:shadow_1}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_2.png}
+    \caption{Awesome Image}
+    \label{fig:shadow_2}
+\end{figure}
+
+%\includegraphics[scale=0.60]{sphere_shadow_figure_2}
+
 Now we use the defintions of $orth_{\vec{u}} \vec{v}$ and $proj_{\vec{u}} \vec{v}$ to simplify equation (\ref{xproperty}).

### 2015-12-14 paper/paper.tex
Moved the figures to be in line with the text, gave them better captions.

@@
 
-\begin{figure}[p]
+\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_1.png}
-    \caption{Awesome Image}
+    \caption{Depicted is the light source, a sphere, and a ray of light tangent to the sphere (labeled as generatrix of cone).}
     \label{fig:shadow_1}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_2.png}
-    \caption{Awesome Image}
+    \caption{This figure includes in addition the perpendicular from the point of tangency to the axis of the cone.
+The distance $L''$ is from the light source to the intersection of this perpendicular with the axis.}
     \label{fig:shadow_2}

### 2015-12-15 paper/paper.tex
Improved a caption.

--- 2015-12-14 paper/paper.tex
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_2.png}
-    \caption{Awesome Image}
+    \caption{This figure includes in addition the orthogonal projection from the point of tangency to the axis of the cone.
+The distance $L''$ is from the light source to the tip of the projection vector.}
     \label{fig:shadow_2}

### 2015-12-15 paper/paper.tex
Removed warnings from Latex source
Additions to the appendix.

@@
 \end{align*}
-Now we find the intersection of this cone with the $z=0$ plane.
+Divide both sides by $k^2$ and substitute $k^2 = r^2 / ( L^2-r^2)$.
+\begin{align*}
+((\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{y})^2)\cdot ( L^2-r^2)/r^2 =  (\vec{d}\cdot\vec{y})^2\\
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 =  (\vec{d}\cdot\vec{y})^2(1 + ( L^2-r^2)/r^2)\\
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 =  (\vec{d}\cdot\vec{y})^2 L^2/r^2
+\end{align*}
+The quantities $\vec{d}\cdot\vec{d}$ and $L^2$ are equal (see figure \ref{fig:shadow_1}).
+\[
+(\vec{y}\cdot\vec{y})( L^2-r^2) =  (\vec{d}\cdot\vec{y})^2
+\]
+Now we find the intersection of this cone with the $z=0$ plane. The condition for intersection is that the third component of $\vec{x}$ be $0$.
 

### 2015-12-15 paper/paper.tex
Fixed the wrong figure numbers

--- 2015-12-15 paper/paper.tex
@@
-\documentclass[letter,10pt]{article}
+\documentclass[10pt]{article}
 \usepackage[utf8]{inputenc}
@@
 Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.
-This allows the bounding of the entire system in a cube, which we can use to define our 0 and our \texttt{MAX\_INT} for each coordinate direction.
+This allows the bounding of the entire system in a cube, which we can use to define our \texttt{0} and our \texttt{MAX\_INT} for each coordinate direction.
 Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.
@@
 \[
-k = {\|orth_{\vec{c}-\vec{l}} (\vec{p} - \vec{l})\|\over \|proj_{\vec{c}-\vec{l}} (\vec{p} - \vec{l})\|} = {r'\over L''} = {r\over L'} = {r\over\sqrt{L^2-r^2}}
+k = \frac{\|orth_{\vec{c}-\vec{l}} (\vec{p} - \vec{l})\|}{\|proj_{\vec{c}-\vec{l}} (\vec{p} - \vec{l})\|} = \frac{r'}{L''} = \frac{r}{L'} = \frac{r}{\sqrt{L^2-r^2}}
 \]
@@
 \begin{align*}
-proj_{\vec{u}} \vec{v} = \vec{u} {\vec{u}\cdot\vec{v}\over \vec{u}\cdot\vec{u}}\\
+proj_{\vec{u}} \vec{v} = \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
 orth_{\vec{u}} \vec{v} = \vec{v} - proj_{\vec{u}} \vec{v}
@@
 \begin{align*}
-\|\vec{y} - \vec{d} {\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}}\|^2 = \|\vec{d} {\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
-\vec{y}\cdot\vec{y}-2\vec{y}\cdot\vec{d} {\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}({\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}})^2 = (\vec{d}\cdot\vec{d}) ({\vec{d}\cdot\vec{y}\over \vec{d}\cdot\vec{d}})^2 \cdot k^2
+\|\vec{y} - \vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 = \|\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
+\vec{y}\cdot\vec{y}-2\vec{y}\cdot\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 = (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
 \end{align*}

### 2015-12-15 paper/paper.tex
Finished draft of appendix A.
Covers everything that should be in the appendix. Variable naming in this derivation is not consistent with the variable naming in rayTracing.py.

--- 2015-12-15 paper/paper.tex
@@
 \DeclareGraphicsExtensions{.pdf,.png,.jpg}
+\usepackage{breqn}
 
@@
 Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.
-This allows the bounding of the entire system in a cube, which we can use to define our \texttt{0} and our \texttt{MAX\_INT} for each coordinate direction.
+This allows the bounding of the entire system in a cube, which we can use to define our 0 and our \texttt{MAX\_INT} for each coordinate direction.
 Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.
@@
 
+We should use bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.
+
 \begin{appendices}
@@
 The first step is to derive the equation of the shadow cone.
-A generic point on the cone will be denoted by $\vec{x}$.
+A generic point on the cone will be denoted by $\vec{s}$.
 The axis of the cone passes through $\vec{l}$ and through $\vec{c}$.
-The defining property of $\vec{x}$ is
-\begin{equation}\label{xproperty}
-\|orth_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| \cdot k
+The defining property of $\vec{s}$ is
+\begin{equation}\label{sproperty}
+\|orth_{\vec{c}-\vec{l}} (\vec{s} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{s} - \vec{l})\| \cdot k
 \end{equation}
 where $k$ is a constant characteristic of the aperture of the cone.
-Equation (\ref{xproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
+Equation (\ref{sproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
 The value of $k$ can be calculated from any point on the cone.
@@
 
-\begin{figure}[h]
+\begin{figure}[h]\label{fig:shadow_1}
     \centering
@@
     \caption{Depicted is the light source, a sphere, and a ray of light tangent to the sphere (labeled as generatrix of cone).}
-    \label{fig:shadow_1}
 \end{figure}
 
-\begin{figure}[h]
+\begin{figure}[h]\label{fig:shadow_2}
     \centering
@@
 The distance $L''$ is from the light source to the tip of the projection vector.}
-    \label{fig:shadow_2}
 \end{figure}
@@
 \begin{align*}
-proj_{\vec{u}} \vec{v} = \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
-orth_{\vec{u}} \vec{v} = \vec{v} - proj_{\vec{u}} \vec{v}
+proj_{\vec{u}} \vec{v} &= \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
+orth_{\vec{u}} \vec{v} &= \vec{v} - proj_{\vec{u}} \vec{v}
 \end{align*}
-Start by squaring both sides of \ref{xproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $y = \vec{x}-\vec{l}$.
+Start by squaring both sides of \ref{sproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $\vec{t} = \vec{s}-\vec{l}$.
 \begin{align*}
-\|\vec{y} - \vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 = \|\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
-\vec{y}\cdot\vec{y}-2\vec{y}\cdot\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 = (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
+\|\vec{t} - \vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}\|^2 &= \|\vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
+\vec{t}\cdot\vec{t}-2\vec{t}\cdot\vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}})^2 &= (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
 \end{align*}
@@
 \begin{align*}
-(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{y})^2+(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2\\
-(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2
+(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{t})^2+(\vec{d}\cdot\vec{t})^2  &= (\vec{d}\cdot\vec{t})^2 \cdot k^2\\
+(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{t})^2 & = (\vec{d}\cdot\vec{t})^2 \cdot k^2
 \end{align*}
+Divide both sides by $k^2$ and substitute $k^2 = r^2 / ( L^2-r^2)$.
+\begin{align*}
+((\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{t})^2)\cdot ( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2\\
+(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2(1 + ( L^2-r^2)/r^2)\\
+(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2 L^2/r^2
+\end{align*}
+The quantities $\vec{d}\cdot\vec{d}$ and $L^2$ are equal (see figure \ref{fig:shadow_1}).
+\begin{equation}\label{coneeqn}
+(\vec{t}\cdot\vec{t})( L^2-r^2) =  (\vec{d}\cdot\vec{t})^2
+\end{equation}
 Now we find the intersection of this cone with the $z=0$ plane.
+Let $x$ and $y$ be coordinates on the shadow plane.
+The condition for intersection is that the third component of $\vec{s}$ be $0$.
+We put $\vec{t}$ and $\vec{d}$ in terms of $x$ and $y$ and substitute into equation \ref{coneeqn}.
+\begin{align*}
+\vec{s}& = \langle x,y,0 \rangle\\
+\vec{d}\cdot\vec{t} &= \vec{d}\cdot\vec{s} - \vec{d}\cdot\vec{l} = d_1x+d_2y - \vec{d}\cdot\vec{l}\\
+\vec{t}\cdot\vec{t} &= \vec{s}\cdot\vec{s} -2\vec{s}\cdot\vec{l} + \vec{l}\cdot\vec{l} = x^2 + y^2-2l_1x-2l_2y+ \vec{l}\cdot\vec{l}\\
+(\vec{d}\cdot\vec{t})^2 &= (d_1x+d_2y - \vec{d}\cdot\vec{l})^2\\
+&= (d_1x+d_2y)^2 - 2(d_1x+d_2y)(\vec{d}\cdot\vec{l}) + (\vec{d}\cdot\vec{l})^2\\
+&= d_1^2x^2+2d_1d_2xy+d_2^2y^2-2d_1x(\vec{d}\cdot\vec{l})-2d_2y(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2\\
+&= d_2^2y^2 + (2d_1d_2x -2d_2(\vec{d}\cdot\vec{l}))y + (d_1^2x^2-2d_1x(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2)\\
+(\vec{t}\cdot\vec{t})( L^2-r^2) &= (x^2 + y^2-2l_1x-2l_2y+ \vec{l}\cdot\vec{l})(L^2-r^2)\\
+&= (L^2-r^2)y^2 + (-2l_2(L^2-r^2))y + (x^2-2l_1x+ \vec{l}\cdot\vec{l})(L^2-r^2)
+\end{align*}
+For the final result we subtract the one side of equation \ref{coneeqn} from the other to obtain:
+\begin{dmath*}
+0 = (\vec{d}\cdot\vec{t})^2 - (\vec{t}\cdot\vec{t})( L^2-r^2)\\
+= (d_2^2 - (L^2-r^2))y^2 + (2d_1d_2x -2d_2(\vec{d}\cdot\vec{l})+2l_2(L^2-r^2))y+(d_1^2x^2-2d_1x(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2 - (x^2-2l_1x+ \vec{l}\cdot\vec{l})(L^2-r^2))
+\end{dmath*}
 

### 2015-12-15 paper/paper.tex

@@
 
-\begin{figure}[h]\label{fig:shadow_1}
+\begin{figure}[h]
     \centering
@@
     \caption{Depicted is the light source, a sphere, and a ray of light tangent to the sphere (labeled as generatrix of cone).}
+    \label{fig:shadow_1}
 \end{figure}
 
-\begin{figure}[h]\label{fig:shadow_2}
+\begin{figure}[h]
     \centering
@@
 The distance $L''$ is from the light source to the tip of the projection vector.}
+    \label{fig:shadow_2}
 \end{figure}

### 2015-12-15 paper/paper.tex
Change the abstract and introduction to better reflect what we did and did not get around to doing.

@@
 This problem was inspired by the artwork of Larry Kagan, a retired
-professor at RPI. The use of classical and real valued genetic
-algorithms to create these objects is demonstrated. A visualization
+professor at RPI. The use of genetic
+algorithms to create these objects is discussed. A visualization
 tool is described that shows how the object would look as a finished
@@
 in the genetic algorithm is the fidelity of the cast shadow to the
-desired result. We use an industry standard algorithm which is often
-applied in image quality analysis to judge the quality of the
-shadows. Having described the means to evaluate the fitness of a
+desired result. We discuss how we evaluate this degree of similarity.
+Having described the means to evaluate the fitness of a
 candidate object, we proceed to explain the recombination techniques
-we used to create each generation of objects from the parent generation.
+we suggest using to create each generation of objects from the parent generation.
 \end{abstract}
@@
 naive sense, the problem  of creating something like Kagan's work is
-a very simple one: create a metal object in the shape of the desired
+a very simple one: one solution is to create a metal object in the shape of the desired
 shadow but scaled down by half, then mount this half way between the
@@
 are its structural stability, its visual chaoticness, and its ability
-to be fabricated using techniques similar to what Kagan uses. We did
-not fully address structural stability in our research. Visual
-chaoticness is achieved by the random nature of the selection process.
-Fabricability is included because it adds another contraint in a
-problem that has few. It is was key consideration when we decided
+to be fabricated using techniques similar to what Kagan uses. Structural stability
+is addressed in section \ref{GA}. Visual
+chaos is achieved by the random nature of the selection process.
+Fabricability is included because it adds another constraint in a
+problem that has few. It was key consideration when we decided
 how to represent these sculptures.
@@
 
-\section{Genetic Algorithm Solutions}
+\section{Genetic Algorithm Solutions} \label{GA}
 We believe approximate solutions to the forward problem may be possible using genetic algorithms for creating ``abstract'' data that fits a given shadow.

### 2015-12-15 paper/paper.tex
Start to fomat the pseudocode as pseudocode in LaTeX.

@@
 \usepackage{breqn}
+\usepackage[]{algorithm2e}
 
@@
 
-input: l = position of light
-	   r = the radius of each sphere and cylinder
-	   spheres = a list of positions of centers of spheres
-	   cylinders = a list of pairs of positions specifying the end faces of the
-					cylinders
-output: pixels = array of black and white pixels indicating on which areas of the
-					wall light is incident
+\begin{dmath*}
+\vec{l} = position of light\\
+\vec{r} = the radius of each sphere and cylinder\\
+spheres = a list of positions of centers of spheres\\
+cylinders = a list of pairs of positions specifying the end faces of the cylinders
+\end{dmath*}
 
-initialize pixels to be uniformly white (illuminated)
+\begin{algorithm}[H]
+\DontPrintSemicolon
+ \KwData{$\vec{l}$ = position of light, $\vec{r}$ = the radius of each sphere and cylinder, $spheres$ = a list of positions of centers of spheres, $cylinders$ = a list of pairs of positions specifying the end faces of the cylinders}
+ \KwResult{$pixels$ = array of black and white pixels indicating on which areas of the wall light is incident}
+ $pixels \leftarrow$ a uniformly white (illuminated) image of the desired size\;
+ \For{each sphere in spheres}{
+  read current\;
+  \eIf{understand}{
+   go to next section\;
+   current section becomes this one\;
+   }{
+   go back to the beginning of current section\;
+  }
+ }
+ \For{each cylinder in cylinders}{
+  \tcp*[h]{similar to the above but cylinders cast different}
+  \tcp*[h]{shaped shadows}\\
+  \tcp*[h]{This case is not handled by our code.}
+ }
+ \caption{first shadow rendering method}
+\end{algorithm}
 
@@
 				the case of elliptical shadows.
-
-for each cylinder in cylinders:
-	//similar to the above but cylinders cast different shaped shadows
-	// ...
-	This case is not handled by our code.
 

### 2015-12-15 paper/paper.tex
Renamed some files for clarity and conformity, updated paper.tex

--- 2015-12-15 paper/paper.tex
@@
 \usepackage[toc,page]{appendix}
-\usepackage{amsmath, amsthm, amssymb}
+\usepackage{amsmath, amsthm, amssymb, amsrefs}
 \usepackage{graphicx}
 \DeclareGraphicsExtensions{.pdf,.png,.jpg}
-\usepackage{breqn}
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{object-render3.png}
+    \caption{A sample render of ten spheres and their shadows.  The light source used was made up of nine point light sources, grouped together very closely.}
+    \label{fig:render_3}
+\end{figure}
+
+To further explain the algorithm behind this method, first let us describe how the object would be represented.
+The method used is most similar to the second representation of objects discussed previously, as spheres and cylinders in a graph layout.
+Each object would consist of a list of nodes or spheres $W$, a list of edges or cylinders $E$, and a radius $r$.
+The nodes in $W$ are described using 3-tuples of floating point numbers, representing their location in Cartesian 3-space, and the edges in $E$ are represented by 2-tuples of indexes, referencing the two nodes in $W$.
+The algorithm iterates over the nodes in $W$, creating spheres of radius $r$.
+Then, it iterates over the edges in $E$, creating cylinders with radius $r$ and endpoints centered on the
+
 For a more updated model that could possibly fix both the issue of slow shadow rendering time and separation of the shadow rendering from the object's rendering, one may instead of legacy OpenGL use OpenGL 3.3 or above, which adds the use of fragment shaders, and can render shadows from point light sources fairly easily.
@@
 This would be the genome for the genetic algorithm.
-The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the methods above.
+The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the shadowmapping methods above.
 Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.
@@
 One may represent an edge between node A and node B as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.
-This allows for easy genetic combinations, with the ability to vary the number of edges in the graph.
+This allows for easy genetic recombinations, with the ability to vary the number of edges in the graph.
 The issue that this does not take care of is the connectivity of the graph.
@@
 
-We should use bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.
-
 \begin{appendices}
@@
 The first step is to derive the equation of the shadow cone.
-A generic point on the cone will be denoted by $\vec{s}$.
+A generic point on the cone will be denoted by $\vec{x}$.
 The axis of the cone passes through $\vec{l}$ and through $\vec{c}$.
-The defining property of $\vec{s}$ is
-\begin{equation}\label{sproperty}
-\|orth_{\vec{c}-\vec{l}} (\vec{s} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{s} - \vec{l})\| \cdot k
+The defining property of $\vec{x}$ is
+\begin{equation}\label{xproperty}
+\|orth_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| \cdot k
 \end{equation}
 where $k$ is a constant characteristic of the aperture of the cone.
-Equation (\ref{sproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
+Equation (\ref{xproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis.
 The value of $k$ can be calculated from any point on the cone.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 \begin{align*}
-proj_{\vec{u}} \vec{v} &= \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
-orth_{\vec{u}} \vec{v} &= \vec{v} - proj_{\vec{u}} \vec{v}
+proj_{\vec{u}} \vec{v} = \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
+orth_{\vec{u}} \vec{v} = \vec{v} - proj_{\vec{u}} \vec{v}
 \end{align*}
-Start by squaring both sides of \ref{sproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $\vec{t} = \vec{s}-\vec{l}$.
+Start by squaring both sides of \ref{xproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $y = \vec{x}-\vec{l}$.
 \begin{align*}
-\|\vec{t} - \vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}\|^2 &= \|\vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
-\vec{t}\cdot\vec{t}-2\vec{t}\cdot\vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}})^2 &= (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
+\|\vec{y} - \vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 = \|\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
+\vec{y}\cdot\vec{y}-2\vec{y}\cdot\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 = (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
 \end{align*}
@@
 \begin{align*}
-(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{t})^2+(\vec{d}\cdot\vec{t})^2  &= (\vec{d}\cdot\vec{t})^2 \cdot k^2\\
-(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{t})^2 & = (\vec{d}\cdot\vec{t})^2 \cdot k^2
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{y})^2+(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2\\
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2
 \end{align*}
@@
 \begin{align*}
-((\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{t})^2)\cdot ( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2\\
-(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2(1 + ( L^2-r^2)/r^2)\\
-(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2 L^2/r^2
+((\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{y})^2)\cdot ( L^2-r^2)/r^2 =  (\vec{d}\cdot\vec{y})^2\\
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 =  (\vec{d}\cdot\vec{y})^2(1 + ( L^2-r^2)/r^2)\\
+(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 =  (\vec{d}\cdot\vec{y})^2 L^2/r^2
 \end{align*}
 The quantities $\vec{d}\cdot\vec{d}$ and $L^2$ are equal (see figure \ref{fig:shadow_1}).
-\begin{equation}\label{coneeqn}
-(\vec{t}\cdot\vec{t})( L^2-r^2) =  (\vec{d}\cdot\vec{t})^2
-\end{equation}
-Now we find the intersection of this cone with the $z=0$ plane.
-Let $x$ and $y$ be coordinates on the shadow plane.
-The condition for intersection is that the third component of $\vec{s}$ be $0$.
-We put $\vec{t}$ and $\vec{d}$ in terms of $x$ and $y$ and substitute into equation \ref{coneeqn}.
-\begin{align*}
-\vec{s}& = \langle x,y,0 \rangle\\
-\vec{d}\cdot\vec{t} &= \vec{d}\cdot\vec{s} - \vec{d}\cdot\vec{l} = d_1x+d_2y - \vec{d}\cdot\vec{l}\\
-\vec{t}\cdot\vec{t} &= \vec{s}\cdot\vec{s} -2\vec{s}\cdot\vec{l} + \vec{l}\cdot\vec{l} = x^2 + y^2-2l_1x-2l_2y+ \vec{l}\cdot\vec{l}\\
-(\vec{d}\cdot\vec{t})^2 &= (d_1x+d_2y - \vec{d}\cdot\vec{l})^2\\
-&= (d_1x+d_2y)^2 - 2(d_1x+d_2y)(\vec{d}\cdot\vec{l}) + (\vec{d}\cdot\vec{l})^2\\
-&= d_1^2x^2+2d_1d_2xy+d_2^2y^2-2d_1x(\vec{d}\cdot\vec{l})-2d_2y(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2\\
-&= d_2^2y^2 + (2d_1d_2x -2d_2(\vec{d}\cdot\vec{l}))y + (d_1^2x^2-2d_1x(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2)\\
-(\vec{t}\cdot\vec{t})( L^2-r^2) &= (x^2 + y^2-2l_1x-2l_2y+ \vec{l}\cdot\vec{l})(L^2-r^2)\\
-&= (L^2-r^2)y^2 + (-2l_2(L^2-r^2))y + (x^2-2l_1x+ \vec{l}\cdot\vec{l})(L^2-r^2)
-\end{align*}
-For the final result we subtract the one side of equation \ref{coneeqn} from the other to obtain:
-\begin{dmath*}
-0 = (\vec{d}\cdot\vec{t})^2 - (\vec{t}\cdot\vec{t})( L^2-r^2)\\
-= (d_2^2 - (L^2-r^2))y^2 + (2d_1d_2x -2d_2(\vec{d}\cdot\vec{l})+2l_2(L^2-r^2))y+(d_1^2x^2-2d_1x(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2 - (x^2-2l_1x+ \vec{l}\cdot\vec{l})(L^2-r^2))
-\end{dmath*}
+\[
+(\vec{y}\cdot\vec{y})( L^2-r^2) =  (\vec{d}\cdot\vec{y})^2
+\]
+Now we find the intersection of this cone with the $z=0$ plane. The condition for intersection is that the third component of $\vec{x}$ be $0$.
 
@@
 
+\begin{bibdiv}
+ \begin{biblist}
+  \bib{Dasgupta}{book}{
+   title={Algorithms},
+   author={Dasgupta, Sanjoy},
+   author={Papadimitriou, Christos},
+   author={Vazirani, Umesh},
+   publisher={McGraw-Hill Education},
+   address={New York},
+   year={2006}  
+  }
+ \end{biblist}
+\end{bibdiv}
+
 \end{document}

### 2015-12-15 render-sample.sh

@@ -1,4 +1,4 @@
 #! /bin/bash
-python make-art-description.py;
+python makeArtDescription.py;
 python rayTracing.py art-description.ad art-description-shadow.jpeg;
-python test1.py art-description.ad art-description-shadow.jpeg;
+python renderObject.py art-description.ad art-description-shadow.jpeg;

### 2015-12-15 paper/paper.tex

--- 2015-12-15 paper/paper.tex
@@
 \usepackage[toc,page]{appendix}
-\usepackage{amsmath, amsthm, amssymb}
+\usepackage{amsmath, amsthm, amssymb, amsrefs}
 \usepackage{graphicx}
@@
 While this may not be optimal, it is the most easily supported method using OpenGL legacy code.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{object-render3.png}
+    \caption{A sample render of ten spheres and their shadows.  The light source used was made up of nine point light sources, grouped together very closely.}
+    \label{fig:render_3}
+\end{figure}
+
+To further explain the algorithm behind this method, first let us describe how the object would be represented.
+The method used is most similar to the second representation of objects discussed previously, as spheres and cylinders in a graph layout.
+Each object would consist of a list of nodes or spheres $W$, a list of edges or cylinders $E$, and a radius $r$.
+The nodes in $W$ are described using 3-tuples of floating point numbers, representing their location in Cartesian 3-space, and the edges in $E$ are represented by 2-tuples of indexes, referencing the two nodes in $W$.
+The algorithm iterates over the nodes in $W$, creating spheres of radius $r$.
+Then, it iterates over the edges in $E$, creating cylinders with radius $r$ and endpoints centered on the
 

### 2015-12-15 paper/paper.tex
Added a new reference article

@@
 The algorithm iterates over the nodes in $W$, creating spheres of radius $r$.
-Then, it iterates over the edges in $E$, creating cylinders with radius $r$ and endpoints centered on the
+Then, it iterates over the edges in $E$, and for each edge $e$ creates cylinders with radius $r$ and endpoints centered on the two nodes in $W$ indexed by $e$.
+This is done using pyOpenGL's
 
@@
   }
+
+  \bib{pyOpenGL}{webpage}{
+   title={PyOpenGL Documentation},
+   url={http://pyopengl.sourceforge.net/documentation/},
+   accessdate={October 2015}
+  }
  \end{biblist}

### 2015-12-15 paper/paper.tex
Did thing.

--- 2015-12-15 paper/paper.tex
@@
 \usepackage[toc,page]{appendix}
-\usepackage{amsmath, amsthm, amssymb, amsrefs}
+\usepackage{amsmath, amsthm, amssymb}
 \usepackage{graphicx}
@@
 \usepackage{breqn}
-\usepackage[]{algorithm2e}
+\usepackage{alltt}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{object-render3.png}
-    \caption{A sample render of ten spheres and their shadows.  The light source used was made up of nine point light sources, grouped together very closely.}
-    \label{fig:render_3}
-\end{figure}
-
-To further explain the algorithm behind this method, first let us describe how the object would be represented.
-The method used is most similar to the second representation of objects discussed previously, as spheres and cylinders in a graph layout.
-Each object would consist of a list of nodes or spheres $W$, a list of edges or cylinders $E$, and a radius $r$.
-The nodes in $W$ are described using 3-tuples of floating point numbers, representing their location in Cartesian 3-space, and the edges in $E$ are represented by 2-tuples of indexes, referencing the two nodes in $W$.
-The algorithm iterates over the nodes in $W$, creating spheres of radius $r$.
-Then, it iterates over the edges in $E$, creating cylinders with radius $r$ and endpoints centered on the
-
 For a more updated model that could possibly fix both the issue of slow shadow rendering time and separation of the shadow rendering from the object's rendering, one may instead of legacy OpenGL use OpenGL 3.3 or above, which adds the use of fragment shaders, and can render shadows from point light sources fairly easily.
@@
 
-\begin{dmath*}
-\vec{l} = position of light\\
-\vec{r} = the radius of each sphere and cylinder\\
-spheres = a list of positions of centers of spheres\\
-cylinders = a list of pairs of positions specifying the end faces of the cylinders
-\end{dmath*}
-
-\begin{algorithm}[H]
-\DontPrintSemicolon
- \KwData{$\vec{l}$ = position of light, $\vec{r}$ = the radius of each sphere and cylinder, $spheres$ = a list of positions of centers of spheres, $cylinders$ = a list of pairs of positions specifying the end faces of the cylinders}
- \KwResult{$pixels$ = array of black and white pixels indicating on which areas of the wall light is incident}
- $pixels \leftarrow$ a uniformly white (illuminated) image of the desired size\;
- \For{each sphere in spheres}{
-  read current\;
-  \eIf{understand}{
-   go to next section\;
-   current section becomes this one\;
-   }{
-   go back to the beginning of current section\;
-  }
- }
- \For{each cylinder in cylinders}{
-  \tcp*[h]{similar to the above but cylinders cast different}
-  \tcp*[h]{shaped shadows}\\
-  \tcp*[h]{This case is not handled by our code.}
- }
- \caption{first shadow rendering method}
-\end{algorithm}
-
+\begin{alltt}
 for each sphere in spheres:
@@
         plane.
-		The equation has the form $A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0$.
+    The equation has the form A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0$.
     Rearrange the equation to have the form $G(x)*y^2 + H(x)*y + I(x) = 0$, where
-				$G$, $H$, and $I$ are 2nd order or lower polynomials in $x$.
-		For each column $c$ in pixels:
-			Find the value of $x$ that corresponds to $c$.
-			Calculate $G(x)$, $H(x)$, and $I(x)$.
-			Use the quadratic formula to find two values of $y$ satisfying the
+        \(G\), \(H$, and \(I\) are 2nd order or lower polynomials in \(x\).
+    For each column \(c\) in pixels:
+      Find the value of \(x\) that corresponds to \(c\).
+      Calculate \(G(x)\), \(H(x)\), and \(I(x)\).
+      Use the quadratic formula to find two values of \(y\) satisfying the
           equation for the boundary of the ellipse.
       Convert the values of y to row coordinates in the image.
-			In column $c$ and all rows $r$ between the two calculated:
+      In column \(c\) and all rows \(r\) between the two calculated:
         color pixels[r][c] black
@@
         the case of elliptical shadows.
+
+\end{alltt}
 
@@
 This would be the genome for the genetic algorithm.
-The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the shadowmapping methods above.
+The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the methods above.
 Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.
@@
 One may represent an edge between node A and node B as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.
-This allows for easy genetic recombinations, with the ability to vary the number of edges in the graph.
+This allows for easy genetic combinations, with the ability to vary the number of edges in the graph.
 The issue that this does not take care of is the connectivity of the graph.
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
     \centering
@@
 
-\begin{bibdiv}
- \begin{biblist}
-  \bib{Dasgupta}{book}{
-   title={Algorithms},
-   author={Dasgupta, Sanjoy},
-   author={Papadimitriou, Christos},
-   author={Vazirani, Umesh},
-   publisher={McGraw-Hill Education},
-   address={New York},
-   year={2006}  
-  }
- \end{biblist}
-\end{bibdiv}
-
 \end{document}

### 2015-12-15 paper/paper.tex
It's compiling now.
w00t.

--- 2015-12-15 paper/paper.tex
@@
         plane.
-    The equation has the form A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0$.
-    Rearrange the equation to have the form $G(x)*y^2 + H(x)*y + I(x) = 0$, where
-        \(G\), \(H$, and \(I\) are 2nd order or lower polynomials in \(x\).
-    For each column \(c\) in pixels:
-      Find the value of \(x\) that corresponds to \(c\).
-      Calculate \(G(x)\), \(H(x)\), and \(I(x)\).
-      Use the quadratic formula to find two values of \(y\) satisfying the
+    The equation has the form A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0.
+    Rearrange the equation to have the form G(x)*y^2 + H(x)*y + I(x) = 0, where
+        G, H, and I are 2nd order or lower polynomials in x.
+    For each column c in pixels:
+      Find the value of x that corresponds to c.
+      Calculate G(x), H(x), and I(x).
+      Use the quadratic formula to find two values of y satisfying the
           equation for the boundary of the ellipse.
       Convert the values of y to row coordinates in the image.
-      In column \(c\) and all rows \(r\) between the two calculated:
+      In column c and all rows r between the two calculated:
         color pixels[r][c] black
@@
 
+for each cylinder in cylinders:
+  //similar to the above but cylinders cast different shaped shadows
+  // ...
+  This case is not handled by our code.
 \end{alltt}

### 2015-12-15 paper/paper.tex
Working on formatting
Fix up the pseudcode a bit.

@@
 \begin{alltt}
-for each sphere in spheres:
-  if the sphere is entirely above the light or below the shadow plane:
+input: \(\vec{l}\) = position of light
+       \(r\) = the radius of each sphere and cylinder
+       \(spheres\) = a list of positions of centers of spheres
+       \(cylinders\) = a list of pairs of positions specifying the end
+              faces of the cylinders
+output: \(pixels\) = array of black and white pixels indicating on
+              which areas of the wall light is incident
+
+initialize \(pixels\) to be uniformly white (illuminated)
+
+for each \(sphere\) in \(spheres\):
+  if the sphere is entirely above the light or below the shadow
+        plane:
     it casts no shadow; continue to the next sphere
@@
         plane.
-    The equation has the form A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0.
-    Rearrange the equation to have the form G(x)*y^2 + H(x)*y + I(x) = 0, where
-        G, H, and I are 2nd order or lower polynomials in x.
+    The equation has the form \(A*x^2 + B*y^2 + C*x*y + D*x + E*y + F = 0\).
+    Rearrange the equation to have the form \(G(x)*y^2 + H(x)*y + I(x) = 0\), where
+        \(G\), \(H\), and \(I\) are 2nd order or lower polynomials in x.
     For each column c in pixels:

### 2015-12-15 paper/paper.tex
Fix referece error and improve pseudocode.

@@
 \documentclass[10pt]{article}
+\usepackage[letterpaper, margin=1in]{geometry}
 \usepackage[utf8]{inputenc}
@@
        \(spheres\) = a list of positions of centers of spheres
-       \(cylinders\) = a list of pairs of positions specifying the end
-              faces of the cylinders
-output: \(pixels\) = array of black and white pixels indicating on
-              which areas of the wall light is incident
+       \(cylinders\) = a list of pairs of positions specifying the end faces of the cylinders
+output: \(pixels\) = array of black and white pixels indicating on which areas of the wall
+              light is incident
 
@@
 for each \(sphere\) in \(spheres\):
-  if the sphere is entirely above the light or below the shadow
-        plane:
+  if the sphere is entirely above the light or below the shadow plane:
     it casts no shadow; continue to the next sphere
@@
         \(G\), \(H\), and \(I\) are 2nd order or lower polynomials in x.
-    For each column c in pixels:
-      Find the value of x that corresponds to c.
-      Calculate G(x), H(x), and I(x).
-      Use the quadratic formula to find two values of y satisfying the
+    For each column \(c\) in pixels:
+      Find the value of \(x\) that corresponds to \(c\).
+      Calculate \(G(x)\), \(H(x)\), and \(I(x)\).
+      Use the quadratic formula to find two values of \(y\) satisfying the
           equation for the boundary of the ellipse.
-      Convert the values of y to row coordinates in the image.
-      In column c and all rows r between the two calculated:
-        color pixels[r][c] black
+      Convert the values of \(y\) to row coordinates in the image.
+      In column \(c\) and all rows \(r\) between the two calculated:
+        color \(pixels[r][c]\) black
   if the sphere intersects the shadow plane:
@@
 
-Now we use the defintions of $orth_{\vec{u}} \vec{v}$ and $proj_{\vec{u}} \vec{v}$ to simplify equation (\ref{xproperty}).
+Now we use the defintions of $orth_{\vec{u}} \vec{v}$ and $proj_{\vec{u}} \vec{v}$ to simplify equation (\ref{sproperty}).
 \begin{align*}

### 2015-12-15 paper/paper.tex
added more references, comment line in renderObject for understanding 57.xxx...

--- 2015-12-15 paper/paper.tex
@@
 \maketitle
+
+\renewcommand{\baselinestretch}{1.5}
 
@@
 Then, it iterates over the edges in $E$, and for each edge $e$ creates cylinders with radius $r$ and endpoints centered on the two nodes in $W$ indexed by $e$.
-This is done using pyOpenGL's
+This is done using pyOpenGL's \texttt{gluSphere()} to draw spheres and \texttt{gluCylinder()} to draw cylinders.
+For more information on the programming in this implementation, see \texttt{renderObject.py}.\ref{gitrepo}
 
@@
   }
-
   \bib{pyOpenGL}{webpage}{
@@
   }
+  \bib{gitrepo}{webpage}{
+   title={Genetic Algorithms for Shadow Sculptures},
+   author={Chris Chappell},
+   author={James Hoctor},
+   url={https://github.com/chappc/Math-Senior-Research},
+   accessdate={December 2015}
+  }
+  \bib{cylindermagic}{webpage}{
+   title={Rendering a Cylinder Between Two Points in OpenGL},
+   author={Parris, Joel},
+   url={http://lifeofaprogrammergeek.blogspot.com/2008/07/rendering-cylinder-between-two-points.html}
+   accessdate={October 2015},
+  }
  \end{biblist}
+
 \end{bibdiv}

### 2015-12-15 renderObject.py

@@ -90,6 +90,7 @@ def cylinder( q, v1, v2, r ):
 
   v = math.sqrt( vx*vx + vy*vy + vz*vz )
   ax = 57.2957795*math.acos( vz/v )
+  # 57.2957795 degrees = 1 radian
   if vz < 0.0:
     ax = -ax
   rx = -vy*vz

### 2015-12-15 paper/paper.tex
Added more to the references section

--- 2015-12-15 paper/paper.tex
@@
 \usepackage{alltt}
+\renewcommand{\baselinestretch}{1.5}
 
@@
 Then, it iterates over the edges in $E$, and for each edge $e$ creates cylinders with radius $r$ and endpoints centered on the two nodes in $W$ indexed by $e$.
-This is done using pyOpenGL's
+This is done using pyOpenGL's \texttt{gluSphere()} to draw spheres and \texttt{gluCylinder()} to draw cylinders.
+For more information on the programming in this implementation, see \texttt{renderObject.py}.\cite{gitrepo}
 
@@
 \end{appendices}
-
+% ftp://ftp.ams.org/ams/amsrefs/amsrdoc.pdf for amsrefs documentation
 \begin{bibdiv}
@@
    url={http://pyopengl.sourceforge.net/documentation/},
-   accessdate={October 2015}
+   accessdate={2015-10}
+  }
+  \bib{gitrepo}{webpage}{
+   title={Genetic Algorithms for Shadow Sculptures},
+   author={Chris Chappell},
+   author={James Hoctor},
+   url={https://github.com/chappc/Math-Senior-Research},
+   accessdate={2015-12}
+  }
+  \bib{cylindermagic}{webpage}{
+   title={Rendering a Cylinder Between Two Points in OpenGL},
+   author={Parris, Joel},
+   url={http://lifeofaprogrammergeek.blogspot.com/2008/07/rendering-cylinder-between-two-points.html},
+   accessdate={2015-10}
   }
  \end{biblist}
+
 \end{bibdiv}

### 2015-12-15 paper/paper.tex
Added a citation.
Added a reference to a citation.

@@
 the mosquito, but created the image of the mosquito with dark
-hard-edged shadows.
+hard-edged shadows. All of these sculptures (with the exception of ``At'') can be viewed on Larry Kagan's website \cite{Kagan}.
 
@@
   }
+  \bib{opengl-tutorial}{webpage}{
+   title={opengl-tutorial},
+   url={http://www.opengl-tutorial.org/},
+   accessdate={November 2015}
+  }
+  \bib{Kagan}{webpage}{
+   title={Larry Kagan Sculpture},
+   url={http://larrykagansculpture.com/},
+   accessdate={October 2015}
+  }
  \end{biblist}

### 2015-12-15 paper/paper.tex
Added a reference to a citation and a reference to another section of the paper.

@@
 The first representation is used when saving a sculpure as a digital file.
-These files can be given as input to the visualization system that is described later in this paper.
-The informal specification of this format can be found in the file `ad file spec.txt' in the git repository submitted with this paper.
+These files can be given as input to the visualization system that is described in section \ref{sculpture_render}.
+The informal specification of this format can be found in the file `ad file spec.txt' in the git repository submitted with this paper \cite{gitrepo}.
 An example file in this format can be generated by the script make-art-description.py, which is also in the git repository.
@@
 
-\section{How We Render Sculptures}
+\section{How We Render Sculptures} \label{sculpture_render}
 Using the method of spheres and cylinders, objects can be graphically modeled and displayed in 3D rendering systems fairly easily.

### 2015-12-15 paper/paper.tex
Fixed a broken reference to the appendix.

@@
 
-The first method is partially implemented in the file `rayTracing.py' on the branch `master' in the submitted git repository.
+The first method is partially implemented in the file `rayTracing.py' on the branch `master' in the submitted git repository \cite{gitrepo}.
 In this method the cast shadow of each sphere and cylinder is computed individually and merged.
@@
 
-Appendix \ref{a} presents a derivation of the equation of the elliptical boundary of the shadow of a sphere in the most typical case.
+Appendix \ref{app:ellipse_derivation} presents a derivation of the equation of the elliptical boundary of the shadow of a sphere in the most typical case.
 
@@
 \begin{appendices}
-\chapter{A} \label{a}
+\section{A} \label{app:ellipse_derivation}
 

### 2015-12-15 paper/paper.tex
Change font used when describing files and git branches.

@@
 
-The first method is partially implemented in the file `rayTracing.py' on the branch `master' in the submitted git repository \cite{gitrepo}.
+The first method is partially implemented in the file \texttt{rayTracing.py} on the branch \texttt{master} in the submitted git repository \cite{gitrepo}.
 In this method the cast shadow of each sphere and cylinder is computed individually and merged.
@@
 In an attempt to render shadows more quickly, another technique was attempted.
-The code for this technique can be found in the newRayTracing branch of the git repository.
+The code for this technique can be found in the \texttt{newRayTracing} branch of the git repository in the file \texttt{rayTracing.py}.
 This second method is a refinement of the following naive method:

### 2015-12-16 paper/paper.tex
Improvements to the paragraphs about the second and third shadow rendering methods.

@@
 Then go through all the spheres and cylinders checking for an intersection.
-At the first intersection, the light intensity at the pixel is decremented and we move on to the next pair of a light and a pixel.
-The motivation behind this method is that there is no step of merging the shadow of each point light source.
-[Consider writing out pseudo code for this.]
+At the first intersection, the light intensity at the pixel is decreased by a fraction of its maximum value and we move on to the next pair of a light and a pixel.
+The motivation behind this method is that it is not necessary to merge the shadow of each point light source, a step that is a substantial time-waster in the first method.
 
 The innermost loop (over all spheres and cylinders) can be accelerated by use of a space partitioning method.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.8\textwidth]{clumping_figure.png}
+    \caption{dfsd}
+    \label{fig:shadow_2}
+\end{figure}
 
@@
 Shadowmapping typically goes on to use this information to shade a rendering of the same scene from another perspective, but we only want an image of the shadow on the wall.
+Therefore we would instead iterate over the pixels in the image and compare the depth of each to the calculated depth at the relevant point in the shadow map.
+If the depths match to within some approximation then that pixel is unshadowed, and inversely, if they differ, the pixel is shadowed.
+That this lighting method is dynamic is advantageous because the shadow will have to be rendered with respect to multiple point light sources.
+Like the first method, this method has the downside of requiring that the shadows from each light be averaged.
 

### 2015-12-16 paper/paper.tex
Add a caption. Fix up the future work section. Add a citation.

@@
     \includegraphics[width=0.8\textwidth]{clumping_figure.png}
-    \caption{dfsd}
+    \caption{The space partitioning scheme used in the second shadow rendering method. The extension to three dimensions is not difficult.}
     \label{fig:shadow_2}
@@
 \section{Future Work}
-What is the best way to use point sources to simulate a diffuse light source? Is there a better way than point light sources to simulate a diffuse source?
+There is a good deal of work left to do before a machine can generate a sculpture in Larry Kagan's style.
+Although we have ideas about how a genetic algorithm could perform this task, once one is implemented there will likely be a good deal of experimentation required before it works well.
+We had also intended to investigate standard metrics for image quality analysis for use judging the accuracy of a shadow in representing an image.
+The time taken to render a single shadow is currently several seconds.
+This cost will be incurred for each fitness evaluation in a genetic algorithm, which strongly suggests that accelerating this process is a prerequisite for experimenting with genetic algorithms.
+There is code, \texttt{shadowMap.py}, in the git repository \cite{gitrepo} that is based heavily on code from opengl-tutorial.org \cite{opengl-tutorial}.
+This code is a very early stage investigation of the use of pyOpenGL to do shadow mapping.
+One current stumbling block with this code is that the C++ code it is based on makes use of a library called OpenGL Mathematics (GLM) which is not available for python \cite{GLM}.
 
-How can the partitioning of space in the second rendering method be improved to allow faster shadow drawing? More equally sized partitions? Adaptively sized partitions?
+There are also a number of interesting questions and developments to investigate, which we list here.
+\begin{itemize}
+  \item What is the best way to use point sources to simulate a diffuse light source? Is there a better way than point light sources to simulate a diffuse source?
 
-The inverse square law is not respected by our shadow drawing algorithms.
-(Does this Include the OpenGL one?) Can this law be incorporated into our algorithms?
+  \item How can the partitioning of space in the second rendering method be improved to allow faster shadow drawing? More equally sized partitions? Adaptively sized partitions?
+
+  \item The inverse square law is not respected by any of our shadow drawing algorithms.
+Can this law be incorporated into our algorithms?
 Is it even necessary to include this or does the human visual system ignore this variation in brightness?
-Mention the image processing technique histogram equalization.
 
-The visualization tool lacks the ability to draw shadows cast on the object by itself. These shadows are not an important component of Kagan's art as far as we can tell.
+  \item The visualization tool lacks the ability to draw shadows cast on the object by itself. These shadows are not an important component of Kagan's art as far as we can tell.
 
-Is it a good idea to combine the two cpu-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?
+  \item Is it a good idea to combine the two cpu-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?
 
-We should use bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.
+  \item In the first shadow rendering method, it would be faster to use Bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.
+\end{itemize}
 
@@
   }
+  \bib{GLM}{webpage}{
+   title={OpenGL Mathematics Library},
+   url={http://glm.g-truc.net/0.9.4/index.html},
+   accessdate={December 2015}
+  }
  \end{biblist}

### 2015-12-16 paper/paper.tex
Added to the description of the second shadow drawing proceedure.

@@
 
-The innermost loop (over all spheres and cylinders) can be accelerated by use of a space partitioning method.
-
 \begin{figure}[ht]
@@
     \includegraphics[width=0.8\textwidth]{clumping_figure.png}
-    \caption{The space partitioning scheme used in the second shadow rendering method. The extension to three dimensions is not difficult.}
-    \label{fig:shadow_2}
+    \caption{The space partitioning scheme used in the second shadow rendering method. The extension to three dimensions is simple.}
+    \label{fig:space_partitioning}
 \end{figure}
 
-The refinement that can be added is simplest to understand if we restrict ourselves temporarily to intersections with spheres.
-For each light source we imagine a pyramid with the region-of-interest of the shadow plane as its base and the light source as its vertex.
-We divide the pyramid into chunks and to each chunk associate all the spheres with a center near the chunk.
-When checking if a line segment inersects any sphere, we can ignore any sphere not in a chunk near the line segment.
-OK, but how do we organize this in practice?
-In our implementation, the pyramid is first divided into horizontal regions...
+The innermost loop (over all spheres and cylinders) can be accelerated by use of a space partitioning method.
+Let us restrict ourselves temporarily to looking for intersections of light paths with spheres.
+By building a data structure before hand that divides the list of spheres based on the locations of their centers, we can disregard most of them during any one intersection check.
+This technique was inspired by $k$-d trees, which are space partitioning data structures used in one solution to the nearest neighbor search problem.
+It turns out that space partitioning methods are very commonly used in ray tracing.
 
-[Mention inspiration from nearest neighbor algorithm]
+For each light source we imagine a rectangular pyramid with the region-of-interest of the shadow plane as its base and the light source as its vertex.
+(See figure \ref{fig:space_partitioning}).
+We divide the pyramid into chunks and to each chunk associate all the spheres with a center in the chunk.
+When checking if a line segment intersects any sphere, we can ignore any sphere not in a chunk near the line segment.
+In particular, we started by dividing the pyramid into four levels.
+At the top level there are no subdivisions as we would have to check the spheres in each division anyway.
+This is because any partition that gets arbitrarity close to the light source can include spheres that shadow the entire shadow plane.
+The next level down is divided into truncated pyramids.
+The top plane of this layer can be though of as divided into squares with a side length equal to $r$ the radius of the spheres.
+A ray of light from the source will pass through one of these squares on its way to the shadow plane unless it is blocked in the previous layer.
+In this second highest layer it can only be blocked by spheres whose centers are beneath that square or any adjacent square.
+Therefore, we check for intersection only with spheres that fall in these partitions.
+In python we implement this by finding the intersection of the light path with the top plane and calculating integer coordinates of the square in which the intersection falls.
+Then from this coordinate pair is calculated a list of adjacent coordinates.
+All of these pairs are used in turn to perform a look-up operation in a python dictionary object each of which results in a list of spheres.
+Only these spheres are checked for intersection.
+The levels beneath that one work the same way.
+The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide th spheres.
+As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere can be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
+This is necessary because the part of the sphere above the plane dividing the levels might block a light path that does not go through an adjacent square.
+The problem is avoided by putting the sphere in the higher level.
+It might seem at first that this trick makes the method no longer a space-partioning method, but it fact it merely means that the partitions are not as drawn in figure \ref{fig:space_partitioning}.
+Instead the top and bottom of a partition is at a distance $r$ below the planes drawn in figure \ref{fig:space_partitioning} (With the exception of the top of the top partition and the bottoms of the bottom partitions).
+
+One way this method could be applied to the cylinders is to first divide the cylinders by size.
+The smaller cylinders can be used with a space-partioning data structure with smaller partitions and the larger ones pun in another space partitioning structure of just put in a list and checked at each iteration.
 

### 2015-12-16 paper/paper.tex
Spelling correction.

@@
 One way this method could be applied to the cylinders is to first divide the cylinders by size.
-The smaller cylinders can be used with a space-partioning data structure with smaller partitions and the larger ones pun in another space partitioning structure of just put in a list and checked at each iteration.
+The smaller cylinders can be used with a space-partioning data structure with smaller partitions and the larger ones put in another space partitioning structure of just put in a list and checked at each iteration.
 

### 2015-12-16 paper/paper.tex
Spelling correction.
Another spelling fix.

@@
 The next level down is divided into truncated pyramids.
-The top plane of this layer can be though of as divided into squares with a side length equal to $r$ the radius of the spheres.
+The top plane of this layer can be thought of as divided into squares with a side length equal to $r$ the radius of the spheres.
 A ray of light from the source will pass through one of these squares on its way to the shadow plane unless it is blocked in the previous layer.
@@
 The levels beneath that one work the same way.
-The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide th spheres.
+The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide the spheres.
 As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere can be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.

### 2015-12-16 paper/paper.tex
Correction in description of second ray tracing method.

@@
 The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide the spheres.
-As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere can be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
+As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere will be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
 This is necessary because the part of the sphere above the plane dividing the levels might block a light path that does not go through an adjacent square.

### 2015-12-16 paper/paper.tex
Corrected an erroneously capitalized word.

@@
 It might seem at first that this trick makes the method no longer a space-partioning method, but it fact it merely means that the partitions are not as drawn in figure \ref{fig:space_partitioning}.
-Instead the top and bottom of a partition is at a distance $r$ below the planes drawn in figure \ref{fig:space_partitioning} (With the exception of the top of the top partition and the bottoms of the bottom partitions).
+Instead the top and bottom of a partition is at a distance $r$ below the planes drawn in figure \ref{fig:space_partitioning} (with the exception of the top of the top partition and the bottoms of the bottom partitions).
 

### 2015-12-16 paper/paper.tex
Change instances of 'shadowmap' and similar to 'shadow map'.

@@
 It should also be very fast to render shadows using openGL.
-Shadowmaps are a standard technique for rendering dynamic shadows.
-In shadowmapping, a scene is rendered from the perspective of a point light source and the resulting ``image'' is colored according to how far each part of the image is from the light.
-Shadowmapping typically goes on to use this information to shade a rendering of the same scene from another perspective, but we only want an image of the shadow on the wall.
+Shadow maps are a standard technique for rendering dynamic shadows.
+In shadow mapping, a scene is rendered from the perspective of a point light source and the resulting ``image'' is colored according to how far each part of the image is from the light.
+Shadow mapping typically goes on to use this information to shade a rendering of the same scene from another perspective, but we only want an image of the shadow on the wall.
 Therefore we would instead iterate over the pixels in the image and compare the depth of each to the calculated depth at the relevant point in the shadow map.
@@
 This would be the genome for the genetic algorithm.
-The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the shadowmapping methods above.
+The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the shadow mapping methods above.
 Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.

### 2015-12-16 paper/paper.tex
Some corrections made to the paper

--- 2015-12-16 paper/paper.tex
@@
 %opening
-\title{Tools and Ideas for the Proceedural Generation of Shadow Sculptures.}
+\title{Tools and Ideas for the Procedural Generation of Shadow Sculptures}
 \author{James Hoctor, Chris Chappell}
@@
 \section*{Introduction}
-This paper addresses the proceedural generation of artwork, if such
+This paper addresses the procedural generation of artwork, if such
 a thing is even possible. We wish to create art in the style of
@@
 that have been bent to shape and welded together to form a whole.
-This assembly is mounted on a wall (painted white) and a spot light
-shines on it to cast a shadow on the wall that forms an image. In a
+This assembly is mounted on a white wall and a spot light
+shines on it to cast a shadow that forms an image on the wall. In a
 naive sense, the problem  of creating something like Kagan's work is
@@
 Fabricability is included because it adds another constraint in a
-problem that has few. It was key consideration when we decided
+problem that has few. It was a key consideration when we decided
 how to represent these sculptures.
@@
 \section{Larry Kagan's Work} \label{kagan}
-In researching this problem we made a brief study of Kagan's work for
+In researching this problem we studied of Kagan's work for
 several reasons. Firstly, we wanted to identify a subset of his work
@@
 
-Each of Larry Kagan's sculptures consists of a metal structure
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.8\textwidth]{mosquito_full.jpg}
+    \caption{Larry Kagan's ``Mosquito I''}
+    \label{fig:mosquito}
+\end{figure}
+
+Each of Kagan's sculptures consists of a metal structure
 composed of bent rods of metal mounted to a wall and illuminated by
@@
 each sculpture, soft egdes of the shadow are noticeable in many of
-these works. In ``Mosquito I'', Kagan uses this effect in a dramatic
+these works. In ``Mosquito I'' \ref{fig:mosquito}, Kagan uses this effect in a dramatic
 way; the image present in the shadow of this sculpture is a mosquito
@@
 the mosquito, but created the image of the mosquito with dark
-hard-edged shadows. All of these sculptures (with the exception of ``At'') can be viewed on Larry Kagan's website \cite{Kagan}.
+hard-edged shadows. All of these sculptures can be viewed on Larry Kagan's website \cite{Kagan}.
 
@@
 Using the method of spheres and cylinders, objects can be graphically modeled and displayed in 3D rendering systems fairly easily.
-One method of doing this that we tested was using OpenGL to build a little engine that could render the object and a small surface such that the object and the shadow could be viewed from multiple angles, as it would look if it were printed.
+One method that we tested was using OpenGL to build a little engine that could render the object and a small surface, such that the object and the shadow could be viewed from multiple angles.
 The shadow could be imported as a texture, using OpenGL's \texttt{GL\_TEXTURE2D} object, and then fitted to the surface that would compose the wall.
@@
 Each object would consist of a list of nodes or spheres $W$, a list of edges or cylinders $E$, and a radius $r$.
-The nodes in $W$ are described using 3-tuples of floating point numbers, representing their location in Cartesian 3-space, and the edges in $E$ are represented by 2-tuples of indexes, referencing the two nodes in $W$.
+The nodes in $W$ are described using 3-tuples of floating point numbers, representing their location in Cartesian 3-space, and the edges in $E$ are represented by 2-tuples of indices, referencing the two nodes in $W$.
 The algorithm iterates over the nodes in $W$, creating spheres of radius $r$.
@@
 
-For a more updated model that could possibly fix both the issue of slow shadow rendering time and separation of the shadow rendering from the object's rendering, one may instead of legacy OpenGL use OpenGL 3.3 or above, which adds the use of fragment shaders, and can render shadows from point light sources fairly easily.
+For a more updated model that could potentially fix both the issue of slow shadow rendering time and separation of the shadow rendering from the object's rendering, one may instead of legacy OpenGL use OpenGL 3.3 or above.
+OpenGL 3.3 adds the use of fragment shaders, and can render shadows from point light sources fairly easily.
 The question that arises in this method, however, is how one would compare two shadows to determine accuracy and fitness of an object.
@@
 There are many video game engines that have builtin methods for shadow generation, with many options and customizable settings for doing so.
-This may be the simplest option for rendering objects, and possibly even would also allow for comparison and heuristic generation.
+This may be the simplest option for rendering objects, and potentially would allow for comparison and heuristic generation.
 Such engines would include Unity, Unreal Engine, and Source 2, just to name a few.
@@
   if the sphere is entirely above the light or below the shadow plane:
-    it casts no shadow; continue to the next sphere
+    it casts no shadow; continue to the next sphere.
   if the top of the sphere is beneath the light and the bottom is above the plane:
@@
       In column \(c\) and all rows \(r\) between the two calculated:
-        color \(pixels[r][c]\) black
+        color \(pixels[r][c]\) black.
   if the sphere intersects the shadow plane:
@@
 
-We did not implement casting the shadows of cylinders by this method, but there is an optimization that deserves mention.
+We did not implement casting the shadows of cylinders by this method, but nonetheless there is an optimization that deserves mention in regards to casting shadows of cylinders.
 It results from the fact that we can assume there is a sphere present at each end of any cylinder.
@@
 The next level down is divided into truncated pyramids.
-The top plane of this layer can be thought of as divided into squares with a side length equal to $r$ the radius of the spheres.
+The top plane of this layer can be though of as divided into squares with a side length equal to $r$ the radius of the spheres.
 A ray of light from the source will pass through one of these squares on its way to the shadow plane unless it is blocked in the previous layer.
@@
 The levels beneath that one work the same way.
-The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide the spheres.
-As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere will be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
+The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide th spheres.
+As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere can be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
 This is necessary because the part of the sphere above the plane dividing the levels might block a light path that does not go through an adjacent square.
@@
 \begin{appendices}
-\section{A} \label{app:ellipse_derivation}
+\section{Shadow Calculations} \label{app:ellipse_derivation}
 
@@
    title={Genetic Algorithms for Shadow Sculptures},
-   author={Chris Chappell},
-   author={James Hoctor},
+   author={Chappell, Chris},
+   author={Hoctor, James},
    url={https://github.com/chappc/Math-Senior-Research},

### 2015-12-16 paper/paper.tex

@@
 The levels beneath that one work the same way.
-The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide th spheres.
-As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere can be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
+The benefit to dividing the lower part of the pyramid further is that lower levels have a greater capacity to divide the spheres.
+As noted in the annotations on figure \ref{fig:space_partitioning}, a sphere will be sorted into a partition in a level higher than its center if it intersects the boundary between the layers.
 This is necessary because the part of the sphere above the plane dividing the levels might block a light path that does not go through an adjacent square.

### 2015-12-16 paper/paper.tex
Changed 'of' to 'or'.

@@
 One way this method could be applied to the cylinders is to first divide the cylinders by size.
-The smaller cylinders can be used with a space-partioning data structure with smaller partitions and the larger ones put in another space partitioning structure of just put in a list and checked at each iteration.
+The smaller cylinders can be used with a space-partioning data structure with smaller partitions and the larger ones put in another space partitioning structure or just put in a list and checked at each iteration.
 

### 2015-12-16 paper/paper.tex
Completed everything before final submition

--- 2015-12-16 paper/paper.tex
@@
 Let us restrict ourselves temporarily to looking for intersections of light paths with spheres.
-By building a data structure before hand that divides the list of spheres based on the locations of their centers, we can disregard most of them during any one intersection check.
+By building a data structure beforehand that divides the list of spheres based on the locations of their centers, we can disregard most of them during any one intersection check.
 This technique was inspired by $k$-d trees, which are space partitioning data structures used in one solution to the nearest neighbor search problem.
@@
 The question then becomes how to represent data for each possible solution to the problem in a genome that allows simple crossovers and mutations.
-Some possible representations are as follows:
+Some possible representations are as follows.
 
 The simplest method of representing nodes and edges would be to keep the number of both constant, moving the nodes around while retaining the same edge mappings for all genomes.
-This way, one could represent each node as an unsigned integer, or an array of three unsigned integers representing the positive incremental distance from some origin point.
+This way one could represent each node as an unsigned integer or an array of three unsigned integers representing the positive incremental distance from some origin point.
 Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.
 This allows the bounding of the entire system in a cube, which we can use to define our \texttt{0} and our \texttt{MAX\_INT} for each coordinate direction.
-Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.
+Next we would create an array of unsigned integers, of length 3 times the number of nodes in the system.
 This would be the genome for the genetic algorithm.
@@
 One may increase the variability of the genetic algorithm by introducing edge mapping to the genome.
-One may represent an edge between node A and node B as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.
-This allows for easy genetic recombinations, with the ability to vary the number of edges in the graph.
+An edge between node A and node B may be represented as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.
+This allows for easy genetic recombinations with the ability to vary the number of edges in the graph.
 The issue that this does not take care of is the connectivity of the graph.
-One would have to either add a penalty to the fitness function which performed a depth-first search on the nodes and penalized unconnected graphs, or alternatively, one could modify the base constructs of the genetic algorithm, making it continue to perform attempts at combining two genomes until it had children that were all fully connected.
-For the first option, the question would become how strongly one would decrease the fitness of an unconnected graph, and whether the number of unconnected pieces should increase that penalty.
+One would have to either add a penalty to the fitness function which performed a depth-first search on the nodes and penalized unconnected graphs; alternatively, one could modify the base constructs of the genetic algorithm to continue to perform attempts at combining two genomes until it had children that were all fully connected.
+For the first option, the question would become how strongly one would decrease the fitness of an unconnected graph and whether the number of unconnected pieces should increase that penalty.
 The second option would allow us to use the same fitness function, but may add extra levels of complexity to the problem due to the required repeated attemps within single generations.
 
-The last important part of the genetic algorithm is a method of performing mutations on genomic data.  Here are some options in consideration:
+The last important part of the genetic algorithm is a method of performing mutations on genomic data.  Here are some options in consideration.
 
 By simply mutating the data in the genome, one would cause one or multiple nodes to be moved to a new location, possibly on the same plane, but possibly not.
-This would be the simplest method of performing a mutation, but depending on how the nodes are laid out, it may mutate a node outside of the acceptable region given for nodes in the given problem.
+This would be the simplest method of performing a mutation, but depending on how the nodes are laid out it may mutate a node outside of the acceptable region given for nodes in the given problem.
 Whether this is allowed or not depends on the situational requirements given.
@@
 
-  \item Is it a good idea to combine the two cpu-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?
+  \item Is it a good idea to combine the two CPU-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?
 
-  \item In the first shadow rendering method, it would be faster to use Bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.
+  \item In the first shadow rendering method, would it be faster to use Bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now?
 \end{itemize}
@@
 \begin{appendices}
-\section{Shadow Calculations} \label{app:ellipse_derivation}
+\section{Ellipse Derivation} \label{app:ellipse_derivation}
 
